---
title: "EC442 Part 3 Problem Set 1: SCF"
authors: ["Alexander Newton","Ed Manuel"]
format: html
editor: visual
code-fold: true
toc: true
---

```{r}
#| code-summary: "Initialise"
#| output: false

library(tidyverse)
library(lodown)
library(survey)
library(mitools)
library(convey)
library(reldist)    
library(REAT)

# if data not already downloaded

# examine all available SCF microdata files
#  scf_cat <-
#    get_catalog( "scf" ,
#                 output_dir = file.path( path.expand( "~" ) , "SCF" ) )
# 
# # # Download 2022 SCF to laptop
#  scf_cat <- subset( scf_cat , year == 2022 )
#  scf_cat <- lodown( "scf" , scf_cat )

scf_imp <- readRDS( file.path( path.expand( "~" ) , "SCF" , "scf 2022.rds" ) )
scf_rw <- readRDS( file.path( path.expand( "~" ) , "SCF" , "scf 2022 rw.rds" ) )



```

```{r}

# I think this basically re-weights the survey and saves as complicated object
scf_design <- 
  svrepdesign( 
    weights = ~wgt , 
    repweights = scf_rw[ , -1 ] , #don't include yy1 column
    data = imputationList( scf_imp ) , 
    scale = 1 ,
    rscales = rep( 1 / 998 , 999 ) ,
    mse = FALSE ,
    type = "other" ,
    combined.weights = TRUE
  )

scf_design$designs <- lapply( scf_design$designs , convey_prep )

scf_MIcombine( with( scf_design , svygini( ~ networth ) ) )

# Approach 2 --------------------------------------------------------------

# This approach uses the "Summary Extract Public Data" provided by the Fed and used in their own publications
# Note this dataset has already constructed the NETWORTH variable we're interested in

# Following here: https://medium.com/@candace_60118/survey-of-consumer-finances-scf-analyzing-weighted-data-in-r-4e4789112c8a
# Also this: https://bookdown.org/jimr1603/Intermediate_R_-_R_for_Survey_Analysis/survey-specific-functions.html 


# Download the files
# download.file("https://www.federalreserve.gov/econres/files/scfp2022excel.zip", "SCFP2022.zip")
# unzip("SCFP2022.zip")
df <- read.csv("SCFP2022.csv")%>%
  select(NETWORTH,WGT,OCCAT1) #WGT here is just the sampling weight, this concatenates each of the implicates with eachother






```

## Constructing a dataset

```{r}

```

## Histograms of wealth

> Make two histograms for: (a) the level of wealth (only plot up to the 95th percentile, otherwise the graph will be hard to read), (b) the logarithm of wealth.

```{r}


## Q1: Make two histograms for: (a) the level of wealth (only plot up to the 95th percentile)

# Find 95th perc
q_95<- wtd.quantile(df$NETWORTH, q=0.95, weight = df$WGT)

# Make hist up to 95th perc
hist1<-df%>%
  filter(NETWORTH<q_95)%>%
  ggplot(mapping = aes(NETWORTH,weight = WGT)) + 
  geom_histogram(binwidth = 100000,
                 color = "white") +
  labs(x = "Net Worth Histogram")

hist1

## (b) the logarithm of wealth.

# Make hist up to 95th perc
hist2<-df%>%
  filter(NETWORTH>0)%>%
  mutate(ln_networth = log(NETWORTH))%>%
  ggplot(mapping = aes(ln_networth,weight = WGT)) + 
  geom_histogram(binwidth = 1,
                 color = "white") +
  labs(x = "Log Net Worth Histogram")

hist2
```

## Wealth shares

> Make a table with the wealth shares of (a) the bottom 50 percent, (b) the next 40 percent, (c) the next 9 percent, (d) the top 1 percent, (e) the top 0.1 percent of households in the sample. At the end, your answers to question 2 and 6 can be summarized in a table like Table 1 below.

```{r}
## Q2: Make a table with the wealth shares of 
# (a) the bottom 50 percent, 
# (b) the next 40 percent, 
# (c) the next 9 percent, 
# (d) the top 1 percent,
# (e) the top 0.1 percent of households in the sample.

# Set up the quantile spacing needed for (a)-(e)
quantiles_q2 = rbind(c(0,0.5),
                  c(0.5,0.9),
                  c(0.9,0.99),
                  c(0.99,1),
                  c(0.999,1))

# Get total wealth to construct shares
totalwealth = sum(df$NETWORTH*df$WGT)

# Find wealth (and entrepeneur) share for each quantile
wealth_share<-list()
ent_share<-list()

for (i in 1:nrow(quantiles_q2)){
quantile_space = quantiles_q2[i,]  
  
wealth_lb<- wtd.quantile(df$NETWORTH, q=quantile_space[1], weight = df$WGT)
wealth_ub<- wtd.quantile(df$NETWORTH, q=quantile_space[2], weight = df$WGT)

wealth_sample<-df%>%
  # extract the relevant sample
  filter(NETWORTH>wealth_lb)%>%
  filter(NETWORTH<wealth_ub)

# Wealth share of that group  
wealth_share [[i]]= sum(wealth_sample$NETWORTH*wealth_sample$WGT) / totalwealth
  
# No. entrepeneurs in that group
ent_sample<-wealth_sample%>%
  filter(OCCAT1==2)

ent_share [[i]]= nrow(ent_sample)/nrow(wealth_sample)
}

# Construct table 
table_shares <- cbind(unlist(wealth_share),unlist(ent_share))
```

## The quantile function

> Plot the quantile function (you will likely have to cut the y-axis to make it readable due to the "dwarves and giants" feature).

```{r}
## Q3: Plot the quantile function

# Choose number of points to fit to 
quantiles_q3 <-(seq(0, 1, by = 0.01))

# Find quantile at each point (using the weights) 
quantile_function<-cbind(quantiles_q3,wtd.quantile(df$NETWORTH, q=quantiles_q3, weight = df$WGT))%>%
  as_tibble%>%
  rename(quantile=quantiles_q3)%>%
  rename(value=V2)

# Create the plot
quantile_plot<-quantile_function%>%
  filter(quantile<0.95)%>%
  filter(quantile>0.05)%>%
  ggplot(aes(x=quantile,y=value))+
  geom_line()+
  theme_bw()

# Alternate way to do all this manually by summing weights
quantile_function2<-df%>%
  arrange((NETWORTH))%>%
  mutate(cum_sum = cumsum(WGT))%>%
  mutate(WGT_TOTAL = sum(WGT))%>%
  mutate(quantile=cum_sum/WGT_TOTAL)%>%
  select(quantile,NETWORTH)

quantile_plot2<-quantile_function2%>%
  filter(quantile<0.95)%>%
  filter(quantile>0.05)%>%
  ggplot(aes(x=quantile,y=NETWORTH))+
  geom_line()+
  theme_bw()

# They look pretty similar 
quantile_plot
quantile_plot2


```

## The Lorenz curve

> Plot the Lorenz curve.

```{r}


## Q4: Plot the Lorenz curve

# That was easy
lorenz(df$NETWORTH, weighting = df$WGT, z = NULL, na.rm = TRUE,
       lcx = "Income share", lcy = "Population share",
       lctitle = "Lorenz curve")

```

## Power law distribution

> For the wealth above the 90th percentile, plot log(1 − F (w)) against log w where F (w) is the cumulative distribution function of wealth w. Figure 1 plots the corresponding graph for the 2007 SCF. If the distribution had an exact Pareto tail, this should be a straight line. The plot you will generate will most likely have the feature that the Pareto relationship breaks down for high enough wealth levels. Why?

```{r}
## Q5:For the wealth above the 90th percentile, plot log(1 − F (w)) against log w 
# where F (w) is the cumulative distribution function of wealth w.

# Extract data for the top 10%
q_90<- wtd.quantile(df$NETWORTH, q=0.90, weight = df$WGT)

wealth_top10<-df%>%
  filter(NETWORTH>q_90)%>%
  # Construct CDF function by summing mass below each point
  arrange((NETWORTH))%>%
  mutate(cum_sum = cumsum(WGT))%>%
  mutate(WGT_TOTAL = sum(WGT))%>%
  mutate(F=cum_sum/WGT_TOTAL)%>%
  # Construct the variables we need
  mutate(F_adj = log(1-F))%>%
  filter(NETWORTH>0)%>%
  mutate(log_w = log(NETWORTH))

# Make plot
plot_q5<-wealth_top10%>%
  filter(F<1)%>%
  ggplot(aes(x=log_w,y=F_adj))+
  geom_line()+
  theme_bw()

plot_q5
```

## Entrepreneurs

> Let's assess what fraction of people in different parts of the wealth distribution are "en- trepreneurs" broadly defined. To this end use the variable X4106. For each of the wealth groups in question 2, calculate the fraction of people who are "entrepreneurs". Add your answer to Table 1.
